Developing model Hamiltonians for single and bilayer graphene using density matrix downfolding

### 2019-09-25 draft.tex
finished!

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{subfigure}
\usepackage[margin=1.0in]{geometry}
\usepackage{float}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{multicol}
\renewcommand{\baselinestretch}{1.5}
\usepackage[utf8]{inputenc}
\usepackage{etoolbox}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}
\usepackage{sectsty}
\sectionfont{\fontsize{12}{15}\selectfont}
\subsectionfont{\fontsize{12}{15}\selectfont}
\usepackage{titlesec}
\titlespacing*{\section}{0pt}{0.2\baselineskip}{0.2\baselineskip}
\titlespacing*{\subsection}{0pt}{0.2\baselineskip}{0.2\baselineskip}
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\usepackage{etoolbox}
\patchcmd{\thebibliography}
  {\settowidth}
  {\setlength{\itemsep}{0pt plus 0.01pt}\settowidth}
  {}{}
\apptocmd{\thebibliography}
  {\small}
  {}{}
\author{Shivesh Pathak}
\title{Developing model Hamiltonians for single and bilayer graphene using density matrix downfolding}

@@

\subsection{Density Matrix Downfolding (DMD)}
The starting point of model development for materials is the \textit{ab initio} Hamiltonian defined on the \textit{ab initio} Hilbert space $\mathcal{H}$
\begin{equation}
\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + hat{H}_\text{ion}(\{R_I\})  
\label{eq:Hab}
\end{equation}
where index $i,j$ correspond to electrons, $I$ to ions, $\hat{H}_\text{ion}$ corresponds to the total ionic energy, and the unit of energy is Hartree (Ha).
The goal is to develop an effective Hamiltonian $\hat{H}_\text{eff}$ defined on a low-energy subspace $\mathcal{LE} \subset \mathcal{H}$ which can accurately reproduce the lowest lying eigenstates and energies of $\hat{H}_\text{ab}$.
$\hat{H}_\text{eff}$ is taken to be a linear operator on $\mathcal{LE}$ and can be expressed as a linear combination of Hermitian operators $\hat{d}_k$, with coupling constants $g_k$ and energy offset $E_0$
\begin{equation}
\hat{H}_\text{eff} = \sum_{k} g_k \hat{d}_k  + E_0.
\label{eq:Heff}
\end{equation}
The process of taking $(\hat{H}_\text{ab}, \mathcal{H}) \rightarrow (\hat{H}_\text{eff}, \mathcal{LE})$ is called downfolding of $\hat{H}_\text{ab}$.

@@

\vspace{-10pt}
\paragraph{Solutions of the model}
The final model in \ref{fig:cuo_regr}c was solved by exact diagonalization.
Shown in Figure \ref{fig:cuo_regr}d is a comparison between the excitation energies calculated from the model and those seen in the latest anion photoelectron spectroscopy measurements \cite{Wu1997}.
The model energies for these eigenstates are denoted by orange circles and agree with experiment up to a maximum deviation of 0.2 eV.
Further, the term symbols assigned in experiment match the term symbols of the model eigenstates, which are constructed by looking at the multiplicity and electronic configuration of each model eigenstate.

@@

\vspace{-10pt}
\paragraph{Next steps}
While PCR was demonstrated here, I also tried feature selection using LASSO and Ridge regression. 
I found a high quality of fit on the sampled data using all three methods, $R^2$ = 0.96 - 0.98, but the model parameters, eigenstates and energies differed significantly between them.
I realized this was because portions of the low-energy space were not accessible with the sampling scheme described above.
%Rephrase
In particular, the DFT calculations were unable to generate two single particle excitations on the ground state, $\text{Cu } 3d_\pi \rightarrow \text{O } 2p_\pi$ and $\text{Cu } 3d_{z^2} \rightarrow \text{O } 2p_\pi$.
Therefore, eigenstates corresponding to these excitations were not constrained by the sampled data, leading to large variations in the eigenvalues between different feature selection schemes.

>>>>

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{subfigure}
\usepackage[margin=1.0in]{geometry}
\usepackage{float}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{multicol}
\renewcommand{\baselinestretch}{1.5}
\usepackage[utf8]{inputenc}
\usepackage{etoolbox}
\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}
\usepackage{sectsty}
\sectionfont{\fontsize{12}{15}\selectfont}
\subsectionfont{\fontsize{12}{15}\selectfont}
\usepackage{titlesec}
\titlespacing*{\section}{0pt}{0.2\baselineskip}{0.2\baselineskip}
\titlespacing*{\subsection}{0pt}{0.2\baselineskip}{0.2\baselineskip}
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\usepackage{etoolbox}
\patchcmd{\thebibliography}
  {\settowidth}
  {\setlength{\itemsep}{0pt plus 0.01pt}\settowidth}
  {}{}
\apptocmd{\thebibliography}
  {\small}
  {}{}
\usepackage{datetime}
\newdate{date}{09}{10}{2019}
\date{\displaydate{date}}
\title{Developing model Hamiltonians for single and bilayer graphene using density matrix downfolding}

@@

\subsection{Density Matrix Downfolding (DMD)}
The starting point of model development for materials is the \textit{ab initio} Hamiltonian defined on the \textit{ab initio} Hilbert space $\mathcal{H}$
\begin{equation}
\hat{H}_\text{ab} = \sum_i -\frac{1}{2} \nabla_i^2 + \sum_{iI}\frac{Z_I}{|r_i - R_I|} + \frac{1}{2} \sum_{ij} \frac{1}{|r_i - r_j|} + \hat{H}_\text{ion}(\{R_I\})  
\label{eq:Hab}
\end{equation}
where index $i,j$ correspond to electrons, $I$ to ions, $\hat{H}_\text{ion}$ corresponds to the total ionic energy, and the unit of energy is Hartree (Ha).
The goal is to develop an effective Hamiltonian $\hat{H}_\text{eff}$ defined on a low-energy subspace $\mathcal{LE} \subset \mathcal{H}$ which can accurately reproduce the lowest lying eigenstates and energies of $\hat{H}_\text{ab}$.
$\hat{H}_\text{eff}$ is taken to be a linear operator on $\mathcal{LE}$ and can be expressed as a linear combination of Hermitian operators $\hat{d}_k$, with coupling constants $g_k$ and energy offset $E_0$
\begin{equation}
\hat{H}_\text{eff} = \sum_{k} g_k \hat{d}_k  + E_0.
\label{eq:Heff}
\end{equation}
The process of taking $(\hat{H}_\text{ab}, \mathcal{H}) \rightarrow (\hat{H}_\text{eff}, \mathcal{LE})$ is called downfolding of $\hat{H}_\text{ab}$.

@@

\vspace{-10pt}
\paragraph{Solutions of the model}
The final model in Figure \ref{fig:cuo_regr}c was solved by exact diagonalization.
Shown in Figure \ref{fig:cuo_regr}d is a comparison between the excitation energies calculated from the model and those seen in the latest anion photoelectron spectroscopy measurements \cite{Wu1997}.
The model energies for these eigenstates are denoted by orange circles and agree with experiment up to a maximum deviation of 0.2 eV.
Further, the term symbols assigned in experiment match the term symbols of the model eigenstates, which are constructed by looking at the multiplicity and electronic configuration of each model eigenstate.

@@

\vspace{-10pt}
\paragraph{Next steps}
While PCR was demonstrated here, I also tried feature selection using LASSO and Ridge regression. 
I found a high quality of fit on the sampled data using all three methods, $R^2$ = 0.96 - 0.98, but the model parameters, eigenstates and energies differed significantly between them.
I realized this was because portions of the low-energy space were not accessible with the sampling scheme described above.
In particular, the DFT method used was unable to generate determinants corresponding to two single particle excitations on the ground state, $\text{Cu } 3d_\pi \rightarrow \text{O } 2p_\pi$ and $\text{Cu } 3d_{z^2} \rightarrow \text{O } 2p_\pi$.
Therefore, eigenstates corresponding to these excitations were not constrained by sampled data, leading to large variations in the eigenvalues between different feature selection schemes.
